```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# During the practical session {#during}

In this practical session, we'll use data from the Wordbank database to explore, graph and model some vocabulary data in Cantonese.

## Step 1: Loading packages
Check the following packages (also called 'libraries') are installed: here, tidyverse, skimr and wordbankr and add some code to the script to load them. If you've worked your way through the RYouWithMe BasicBasics units, you'll be familiar with the first three. The fourth we'll talk more about in step 2.

> **TASK 1A**: Check the packages are installed. **HINT**: Look at the 'Packages' tab in the bottom right panel. If they have not been installed previously, use the code snippet below to install them.

```{r eval=FALSE}
install.packages()
```

> **TASK 1B**: Load the packages, using the code snippet below

```{r eval=FALSE}
library()
```

## Step 2: Read in the data
We will be working with some 'Wordbank' data. Wordbank is a large open database of children's vocabulary development. More commonly, you'll be accessing data by reading in a file (such as a .csv file). However, the Workbank data can be accessed from R via the wordbankr package. You can read more about Wordbank on their [**website**](http://wordbank.stanford.edu) and in their free [**online book**](https://wordbank-book.stanford.edu).

>**TASK 2A**: Let's check the vignette of the wordbankr package to learn a bit more about the package and accessing the data. If you have successfully installed the wordbankr package, you can type ??wordbankr in the console. A link to the vignette should appear in the 'Help' tab in the bottom right window. Click on the link, have a look at the vignette and try to work out what the `get_administration_data()` function does? How is it different from the `get_item_data()` function? If this doesn't work for you, you can also access the vignette [**here**](https://langcog.github.io/wordbankr/articles/wordbankr.html).

Now let's read in data for words and sentences in Cantonese.

>**TASK 2B**: Use the code snippet below to read in data for WS (words and sentences) form in Cantonese. Assign the data to a dataframe called 'voc_cantonese'.

```{r eval=FALSE}
get_adminstration_data(language = "", form = "")
```

## Step 3: Exploring the data
Now let's get a feel for these data. Use one or more of the following functions to have a look: `dim()`, `str()`, `glimpse()`, `head()`, `tail()`, `skim()` and `summary()`.

>**TASK 3**: Answer the following questions: How many observations are there for how many variables? Which of these variables numeric? Which of these variables contain missing values ('NA's)? What type of variable is 'longitudinal'? What is the minimum child age? Which of these functions do you find most informative and why?

```{r eval=FALSE}
dim(voc_cantonese)
str(voc_cantonese)
glimpse(voc_cantonese)
head(voc_cantonese)
tail(voc_cantonese)
skim(voc_cantonese)
summary(voc_cantonese)
```

## Step 4: Select relevant variables
We don't need all the variables that are currently in the data frame. Let's use the `select()` function (from the dplyr package, part of the tidyverse) to keep only those that we do need.

>**TASK 4A**: Use the bit of code below to keep the following variables: age, language, sex, production and assign these to a new dataframe called 'voc_cantonese_select'. Check you understand what each bit in the code does. Remember, you can type ?select in the console to get more information on the `select()` function.

```{r eval=FALSE}
voc_cantonese_select <- select(voc_cantonese, age, language, sex, production)
```

Sometimes it is helpful to save the smaller dataframe to a csv file. This will allow you to just read in the relevant bits in future, without having to read in the complete data (which can be time-consuming if the complete data contains thousands of rows and lots of variables). We can do both of these steps in one go, using what is called a 'pipe' using the code below. Essentially, we are taking the dataframe with Cantonese vocabulary data (voc_cantonese), and then select relevant variables, and then write those to a csv-file.

>**TASK 4B**: Use the code snippet below to do this. Again, check you understand what each bit in the code does and please do ask if you're not sure.

```{r eval=FALSE}
voc_cantonese_select <- voc_cantonese %>%
  select(age, language, sex, production) %>%
  write_csv(here('voc_cantonese_select.csv'))
```

If, for some reason, you weren't able to access the wordbankr database, you can download the data here [voc_cantonese_select.csv](files/voc_cantonese_select.csv). Put this file in your working directory (where 'here' points to).

>**TASK 4C**: Just to practise reading in data from a .csv file, use the code snippet below to do so.

```{r eval=FALSE}
voc_cantonese_select <- read_csv(here('voc_cantonese_select.csv'))
```

## Step 5: Visualise the data
We'll start with making some histograms to get an idea of the distribution of some of the variables.

>**TASK 5A**: Use the code snippet below to make a histogram for 'age'.

```{r eval=FALSE}
ggplot(voc_cantonese_select, aes(x = age)) + #Tell R what data to use in the first argument, then specify variable you want to plot in x = 
  geom_histogram() +                         #The geom 'geom_histrogram' tells R how to plot the data
  labs(x = "Age (in months)") +              #labs() can be used to label axes properly
  theme_bw()                                 #There are lots of different 'themes' in ggplot that give you options in terms of the look of the plot
```

>**TASK 5B**: Using the code snippet above as a template, now add some code to the script to make a histogram for the productive vocabulary measure 

>**TASK 5C**: Look at the histograms. What do they tell you about the variables age and productive vocabulary?

Let's now plot the productive vocabulary data as a function of age in a scatterplot.

>**TASK 5D**: Use the code snippet below to create a scatterplot with productive vocabulary and age.

```{r eval=FALSE}
ggplot(voc_cantonese_select, aes(x = age, y = production)) +
  geom_point() +
  theme_bw() +
  labs(x = "Age (in months)", y = "Productive vocabulary")
```

In this plot, each dot represents the data from one child. Some are plotted on top of each other (having the same or very similar scores). We can use a different geom to help with that.

>**TASK 5E**: Use the code snippet below to do so. To find out more on what the `geom_jitter()` function does and what further options it gives you, type ?geom_jitter in the console. 

```{r eval=FALSE}
ggplot(voc_cantonese_select, aes(x = age, y = production)) +
  geom_jitter() +
  theme_bw() +
  labs(x = "Age (in months)", y = "Productive vocabulary")
```

This plot does make more points visible, but there is still quite a few overlapping points. We can use the 'alpha' aesthetic to improve the plot further.

>**TASK 5F**: Have a look at [**this website**](https://ggplot2.tidyverse.org/reference/aes_colour_fill_alpha.html) to work out what the 'alpha' aesthetic does. What does it do?

Let's get a better idea of this by specifying a different 'alpha' value as an argument of the geom_jitter() function.

>**TASK 5G**: Specify a value of alpha as an argument to the `geom_jitter()` function to make the data points more transparent.

>**TASK 5H**: Have a look at the new plot. What do the darker areas of the plot represent?

So far we have plotted the data for boys and girls together. But there is evidence that vocabulary development for boys and girls differs and that variable is present in the dataset. We can plot data for boys and girls separately by choosing different colours.

>**TASK 5I**: Use the code snippet below to use different colours for boys and girls.

```{r eval=FALSE}
ggplot(voc_cantonese_select, aes(x = age, y = production, colour = sex)) +
  geom_jitter(alpha = 0.3) +
  theme_bw() +
  labs(x = "Age (in months)", y = "Productive vocabulary") 
```

Finally, we can add lines of best fit to summarise the relationships between age and productive vocabulary for boys and girls.
>**TASK 5J**: Use the code snippet below to add lines of best fit. What do you conclude from the plot?

```{r eval=FALSE}
ggplot(voc_cantonese_select, aes(x = age, y = production, colour = sex)) +
  geom_jitter(alpha = 0.3) +
  geom_smooth(method = 'lm', se = TRUE) +
  theme_bw() +
  labs(x = "Age (in months)", y = "Productive vocabulary")
```

## Step 6: Model the data
First, it is useful to compute some descriptive statistics, and we'll do so for boys and girls separately, using the `summarise()` function.

>**TASK 6A**: Use the code snippet below to calculate descriptive statistics. Check you understand what each part of the code does.

```{r eval=FALSE}
descriptives <- voc_cantonese_select %>%
  summarise(mean_voc = mean(production, na.rm = TRUE),
            sd_voc = sd(production, na.rm = TRUE),
            min_voc = min(production, na.rm = TRUE),
            max_voc = max(production, na.rm = TRUE),
            mean_age = mean(age, na.rm = TRUE),
            sd_age = sd(age, na.rm = TRUE),
            min_age = min(age, na.rm = TRUE),
            max_age = max(age, na.rm = TRUE))
descriptives
```

Now, we will fit a regression model to test the hypothesis that age and sex account for variation in the number of words that a child produces. To keep it simple, we'll use linear regression, although there are reasons to believe that is a simplification (the relationship might not be linear, but this is just to give you an idea). To specify the regression model we use the `lm()` function. This takes as arguments the outcome or dependent variable y (production) as a function of the predictor or independent variables x1 (age) and x2 (sex), as well as the data to be used. We assign this output to a dataframe called 'mod' (for model). The summary() function summarises the main results of the model; we've assigned that summary to a dataframe called 'mod_summary'. To see the summary in the console, we have to call it.

>**TASK 6B**: Use the code snippet below to fit a linear model.

```{r eval=FALSE}
mod <- lm(production ~ age + sex, data = voc_cantonese_select) 
mod_summary <- summary(mod)
mod_summary
```

When we call 'mod_summary', R writes the content of that dataframe to the console. It tells us that the overall model is significant (p-value: <2.2e-16), and that together, the predictors (age and sex) account for 52.2% of variance in the number of words children produce (Adjusted R-squared = 0.522*100). Under 'coefficients' you can see that both predictors (age and sex) are significant (values under 'Pr(>|t|) are smaller than 0.05). You can also see that the estimate for age is positive, indicating that with increasing age, productive vocabulary increases. The negative value for the estimate for sex indicates that boys ('Male' category) score lower than girls (the reference category, in this case 'Female').

## Step 7: Write up the results

You could write this up as follows: "A linear regression was performed with productive vocabulary (*M* = 273, *SD* = 220) as the outcome variable and age (*M* = 23.2, *SD* = 4.2) and sex (485 girls, 502 boys) as the predictor variables. The results of the regression indicated that the model significantly predicted productive vocabulary (*F*(2, 984) = 539.4, *p* < .001, adjusted R2 = 0.52), accounting for 52% of the variance. Age was a significant positive predictor (*beta* = 37.29, *p* < .001): as age increased, productive vocabulary increased. Boys scored lower than girls (beta = -49.92, p < .001)."

This follows APA guidelines which are widely used within psychology and psycholinguistics. The [**Purdue Writing lab website**](https://owl.purdue.edu/owl/research_and_citation/apa6_style/apa_formatting_and_style_guide/statistics_in_apa.html)) is helpful for guidance on punctuating statistics). Journals in linguistics might use different guidelines.

# Answers
## Step 1: Loading packages
**TASK 1B**
library(here)
library(tidyverse)
library(skimr)
library(wordbankr)

## Step 2: Read in the data
**TASK 2A**
The `get_administration_data()` function gives by-administration information. You can specify language and/or form.
The `get_item_data()` function gives by-item information. Again, you can specify language and/or form.

**TASK 2B**
```{r eval=FALSE}
voc_cantonese <- get_adminstration_data(language = "Cantonese", form = "WS"")
```
## Step 3: Exploring the data

**TASK 3**
* There are 987 observations for 15 variables.
* The variables data_id, age, comprehension, and production are numeric.
* The following variables contain missing values: zygosity, birth_order, ethnicity, mom_ed.
* The variable 'longitudinal' is a logical variable that can have the values 'TRUE' or 'FALSE'.
* The minimum child age is 16 months.
* Personally, I find the skim() function most informative as it groups different types of variables (character, logical, numeric etc.) and provides useful overviews (missing values, some summary statistics). I do however, also always use the head() and tail() functions to get a better feeling for the data. 

## Step 5: Visualise the data
**TASK 5B**
```{r eval=FALSE}
ggplot(voc_cantonese_select, aes(x = production)) + 
  geom_histogram() +
  labs(x = "Productive vocabulary") +
  theme_bw()
```

**TASK 5C**
There are roughly similar numbers of children across ages between 16 and 30 months. The vocabulary measure is highly skewed, with more children producing 50 words or less.

**TASK 5F**
The value of the alpha aesthetic determines the opacity (or transparency) of the individual data points.

**TASK 5G**
```{r eval=FALSE}
ggplot(voc_cantonese_select, aes(x = age, y = production)) +
  geom_jitter(alpha = 0.3) +
  theme_bw() +
  labs(x = "Age (in months)", y = "Productive vocabulary") 
```

**TASK 5H**
Darker areas of the plot reflect more overlapping data points in those areas.

**TASK 5J**
From the plot with lines of best fit, we can see that, on average, girls produce slightly more words than boys.
